Comparing probabilistic predictive models applied to football

### 2015-07-07 template.tex
Alteracoes.

%  LaTeX support: latex@mdpi.com\textbf{}
%  In case you need support, please attach any log files that you could have, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

@@

\usepackage{changes}
\definechangesauthor[name={Rafael}, color=blue]{raf}
\setremarkmarkup{(#2)}
\definechangesauthor[name={Danilo}, color= violet]{danilo}
\setremarkmarkup{(#2)}
\definechangesauthor[name={Luis}, color= magenta]{luis}
\setremarkmarkup{(#2)}

@@

	This paper is organized as follows.
	Section \ref{sec::experimental} describes the proposed models, Section \ref{sec::scoring} presents the scoring rules and criteria used in this work to classify the models.
	Section \ref{sec::results} reports the performance of the models predicting 190 matches comprising the second half of the Brazilian championship of first division.
	In Section \ref{sec::remarks} we discuss the results and close with proposals for future research.

@@

	In this section, we present the basic assumptions adopted by the benchmark models and 
	by the three proposed models. The predictions of the
	two benchmark models were published before each matchday at
	\url{http://chancedegol.uol.com.br} and
	\url{http://www.previsaoesportiva.com.br}.

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\subsection{Benchmark Models}
	\label{sec::Benchmark}

@@

	We adopt a Bayesian approach to calculate the prediction probabilities of an upcoming match of \added[id=danilo]{a} team \replaced[id=danilo]{$t$}{$T$} \deleted[id = danilo]{($A$ or $B$)} based on its past performance, i.e\added[id=danilo]{.}, the number of matches it has won, tied and lost.

	Let us consider the outcome of \replaced[id=danilo]{a given}{each} match of team \replaced[id=danilo]{$t$}{$T$} as a categorical random quantity $X$ that may assume only the values $1$ \added[id=danilo]{(if team $t$ wins)}, $2$ \added[id=danilo]{(if a draw occurs)}, $3$ \added[id=danilo]{(if team $t$ loses)}, \deleted[id=danilo]{where $1$ means win ($W$), $2$ draw ($D$) and $3$ loss ($L$)}.
	Denoting by $\theta_1, \theta_2$ and $\theta_3$ ($\theta_3 = 1-\theta_1 - \theta_2$), the probabilities of win, draw and loss, respectively, the probability mass function of $X$ is
	\[
	P(X=x | \boldsymbol{\theta}) = \theta_1^{\mathbb{I}_{\{1\}}(x)}
	\theta_2^{\mathbb{I}_{\{2\}}(x)}(1 - \theta_1 -
	\theta_2)^{{\mathbb{I}_{\{3\}}}(x)}, \qquad x \in \mathcal{X},
	\]

	\noindent
	where $\mathcal{X}=\{1,2,3\}$ is the support of $X$,
	$\mathbb{I}_{\{i\}}(x)$ is the indicator function that assumes the
	value 1 if $x$ equals $i$ and 0 otherwise, and $\boldsymbol{\theta}
	= (\theta_1, \theta_2)$ \replaced[id=danilo]{belongs}{belogns} to $\boldsymbol{\Theta} =
	\{(\theta_1,\theta_2)\in [0,1]^2: \theta_1+\theta_2 \leq 1 \}$, the 2-simplex.

	Assuming that \added[id=danilo]{the outcomes from} $n$ matches of team \replaced[id=danilo]{$t$}{$T$}, given $\boldsymbol{\theta}$, are i.i.d. quantities with the above categorical distribution, and denoting by $M_1$, $M_2$ and $M_3$ the number of matches won, tied or lost by team \replaced[id=danilo]{$t$}{$T$}, the random vector $(M_1, M_2, M_2)$ has multinomial distribution with parameters $n$ and $\boldsymbol{\theta}$ given by
	\[
	P(M_1=n_1,M_2=n_2,M_3=n_3| n, \boldsymbol{\theta})=
	\frac{n!}{n_1!n_2!(n-n_1-n_2)!}\theta_1^{n_1}\theta_2^{n_2}(1-\theta_1-\theta_2)^{n-n_1-n_2},
	\]
	\noindent 
	where $n_1 + n_2 + n_3 = n$.

	Our goal is to compute the predictive posterior distribution of the
	upcoming match, $X_{n+1}$, that is,
	$P(X_{n+1}=x|M_1=n_1,M_2=n_2,M_3=n_3)$, $x\in\mathcal{X}$. Using
	Bayesian inference this probability can be easily computed as
	follows. Suppose that $\boldsymbol{\theta}$ has Dirichlet prior distribution
	with parameter $(\alpha_1,\alpha_2,\alpha_3)$, denoted
	$\mathcal{D}(\alpha_1,\alpha_2,\alpha_3)$, with density function
	\[
	\pi(\boldsymbol{\theta}|\replaced[id=danilo]{\boldsymbol{\alpha}}{\alpha})=\frac{\Gamma(\alpha_1+\alpha_2+\alpha_3)}{\Gamma(\alpha_1)\Gamma(\alpha_2)\Gamma(\alpha_3)}\theta_1^{\alpha_1-1}\theta_2^{\alpha_2-1}(1-\theta_1-\theta_2)^{\alpha_3-1}
	\]
	\noindent for $\alpha_1$, $\alpha_2$, $\alpha_3 > 0$, then the
	posterior distribution of $\boldsymbol{\theta}$ is
	$\mathcal{D}(n_1+\alpha_1,n_2+\alpha_2,n_3+\alpha_3)$. Thus, the
	predictive distribution of $X_{n + 1}$ is given by the
	integral
	$$
	P(X_{n + 1} = x | M_1 = n_1, M_2 = n_2, M_3 = n_3) = \int_{\boldsymbol{\Theta}} P(X_{n
		+ 1} = x | \boldsymbol{\theta}) \pi(\boldsymbol{\theta} | M_1 = n_1, M_2 = n_2, M_3
	= n_3) d\boldsymbol{\theta},
	$$
	which leads to the following probabilities of win, tie and loss:
	\begin{align*}
	P(X_{n+1} = 1 | M_1=n_1,M_2=n_2,M_3=n_3) &=
	\frac{n_1+\alpha_1}{n+\replaced[id=danilo]{\alpha_{\bullet}}{\alpha}}\\
	& \\
	P(X_{n+1} = 2 | M_1=n_1,M_2=n_2,M_3=n_3) &=
	\frac{n_2+\alpha_1}{n+\replaced[id=danilo]{\alpha_{\bullet}}{\alpha}} \\
	& \\
	P(X_{n+1} = 1 | M_1=n_1, M_2=n_2, M_3=n_3) &=
	\frac{n_3+\alpha_1}{n+\replaced[id=danilo]{\alpha_{\bullet}}{\alpha}}
	\end{align*}
	\noindent where $\replaced[id=danilo]{\alpha_{\bullet}}{\alpha} =\alpha_1+\alpha_2+\alpha_3$.

	 \replaced[id=danilo]{In this work, we are interested in checking the predictive performance of the Multinomial-Dirichlet model for the second round of the championship given the results from the first round. In order to do so, we use full time results from the first round as historical data for construction of the Dirichlet prior. Predictions are carried out by using a Bayes information updating mechanism: first, we assign a uniform prior on  $\boldsymbol{\theta}$ over the 2-simplex, i.e. $\mathcal{D}(1, 1, 1)$, and combine this prior with the first-round data, obtaining a posterior Dirichlet distribution through conjugacy that represents the information about $\boldsymbol{\theta}$ up to the first round; then, the posterior of the first round becomes the prior for the second round, which will be combined with the second round matches up to date in order to yield predictive distributions for future games.}{In this work we used the uniform prior on the 2-simplex, i.e. $\mathcal{D}(1,1,1)$. This is not a restrictive prior because, since we predicted only the second round of the championship, the counts of the first round were used to update this prior---thanks to the conjugacy property of Dirichlet priors---and we may say that the actual prior we used is therefore the number of matches won, tied or lost by each team added by one. If we had started our predictions in the first round, this prior would not have given a good predictive performance, but in this case we could have used team rankings such the ones provided by federations and leagues to assess a more realistic prior for each team instead.}\footnote{On the other hand, the two black-box models make predictions from a frequentist perspective; BB1 uses scores of every match in the previous two years, while BB2 uses the match scores of the current championship only.}

	Before we explain and illustrate the proposed models with an example, we make two further remarks.
	The first one is that we will separately consider home and away games for each team. This will allow us to take into account the different performances under those conditions.
	The second remark is that the prediction of an upcoming match between \added[id=danilo]{two teams} $A$ and $B$ can be done using
	the past performance of either team $A$ or $B$.
	To solve this problem, a compromise between the two predictive distributions was adopted. 
	The next subsections \ref{sec::Mn_Dir1}, \ref{sec::Mn_Dir2} and \ref{sec::Mn_Dir3} present three different possible solutions.

@@

	Thus, considering $X_{n + 1}$ the random outcome of this match with
	respect to the home team (Gr\^{e}mio), the predictive probabilities
	of $X_{n + 1}$ is obtained by equally weighting the two predictive
	distributions, resulting
	\begin{align*}
	P(X_{n+1}=1|{\bf h}, {\bf a}) &=
	\frac{1}{2}\left(\frac{h_1+\alpha_1}{\replaced[id=danilo]{h_{\bullet}+\alpha_{\bullet}}{h+\alpha}}\right)+\frac{1}{2}\left(\frac{a_3+\alpha_3}{\replaced[id=danilo]{a_{\bullet}+\alpha_{\bullet}}{a+\alpha}}\right)=0.5
	\\
	P(X_{n+1}=2|{\bf h}, {\bf a}) &=
	\frac{1}{2}\left(\frac{h_2+\alpha_2}{\replaced[id=danilo]{h_{\bullet}+\alpha_{\bullet}}{h+\alpha}}\right)+\frac{1}{2}\left(\frac{a_2+\alpha_2}{\replaced[id=danilo]{a_{\bullet}+\alpha_{\bullet}}{a+\alpha}}\right)\simeq0.2917, \\
	P(X_{n+1}=3|{\bf h}, {\bf a}) &= \frac{1}{2}
	\left(\frac{h_3+\alpha_3}{\replaced[id=danilo]{h_{\bullet}+\alpha_{\bullet}}{h+\alpha}}\right)+\frac{1}{2}\left(\frac{a_1+\alpha_1}{\replaced[id=danilo]{a_{\bullet}+\alpha_{\bullet}}{a+\alpha}}\right)\simeq0.2083.
	\end{align*}
	\noindent where $\replaced[id=danilo]{h_{\bullet}}{h}=h_1+h_2+h_3$ and $\replaced[id=danilo]{a_{\bullet}}{a}=a_1+a_2+a_3$.

@@

	The model $Mn-Dir_2$ is very similar to model $Mn-Dir_1$, but the
	weights used to mix the posterior distributions are not the same.
	Model $Mn-Dir_2$ weights the predictive distributions according to
	the team ranks just before the match is played. In the illustrative
	example, we know that after 19 matchdays, Gr\^emio was the $6^{th}$
	place and Atl\'etico-PR the $11^{th}$. Therefore, we shall consider
	the weights
	\[w_1 = \frac{1/6}{1/6+1/11}=0.647 ~ ~ ~ ~ ~\text{and} ~ ~ ~ w_2 = \frac{1/11}{1/6+1/11}=0.353\]
	\noindent for the Gr\^emio and Atl\'etico predicitive distributions,
	respectively. Thus, the winning probability for Gr\^emio is
	calculated as
	\[P(X_{n+1}= 1|{\bf h}, {\bf a})=0.647\left(\frac{h_1+\alpha_1}{\replaced[id=danilo]{h_{\bullet}+\alpha_{\bullet}}{h+\alpha}}\right)+0.353\left(\frac{a_3+\alpha_3}{\replaced[id=danilo]{a_{\bullet}+\alpha_{\bullet}}{a+\alpha}}\right)=0.525,\]
	\noindent and the other probabilities are computed similarly.

	\subsection{Model Three: Multinomial-Dirichlet 3}
	\label{sec::Mn_Dir3}

@@

	Considering the example above, the combined vector ${\bf c}$ is given by ${\bf c}=(c_1,c_2,c_3)=(10,5,3)$, which leads to a posterior $\mathcal{D}(11,6,4)$ when considering the prior $\mathcal{D}(1,1,1)$.
	The probability that Gr\^emio wins is therefore	
	\[P(X_{n+1}=1|{\bf c})=\frac{c_1+\alpha_1}{\replaced[id=danilo]{c_{\bullet}+\alpha_{\bullet}}{c+\alpha}}\simeq0.524
	\]

	\noindent
	where $\replaced[id=danilo]{c_{\bullet}}{c}=c_1+c_2+c_3$. The probabilities of drawing and losing are computed similarly.

@@

	How can we fairly rank the various predictive models?
	A way of doing this is by using proper scoring rules, where
	the score may be interpreted as a numerical measure of how inaccurate was a given probabilistic prediction.

	Formally, let $X$ be a random variable
	taking values in $\mathcal{X}=\{1,2,3\}$ indicating
	the outcome of the match, with 1 standing for home win, 2 for draw and 3 for away win.  Moreover, let $P=(P_1,P_2,P_3)$ denote one's probabilistic prediction
	about $X$, i.e., $P$ lies in the 2-simplex set $\Delta_2=\{(p_1,p_2,p_3):p_1+p_2+p_3=1, \ p_1,p_2,p_3\geq0\}$ (see Figure
	\ref{fig:simplex}). 	
	A scoring rule is a function 
	that assigns a real number (score) $S(x,P)$ to each $x \in \mathcal{X}$
	and $P \in \Delta_2$ 
	such that 
	for any given $x$ in $\mathcal{X}$, the score  $S(x,P)$ is minimized when $P$ is 
	$(1,0,0)$, $(0,1,0)$ or $(0,0,1)$ depending if $x$ is 1, 2 or 3, respectively.
	The score $S(x,P)$ can be thought as
	a penalty to be paid when one assigns the 
	probabilistic prediction $P$ and outcome
	$x$ occurs. Also, the ``best"\footnote{A scoring rule can also be defined to be such that a \emph{large} value of the score indicates good forecasts.} possible score (i.e., the smallest score value) is achieved when the probabilistic prediction for the outcome of the game is perfect.
	%Most of the scoring rules also exhibit an arbitrary normalizing feature such as $S(x,P_X)=0$ for each $x$.

	Although many functions can satisfy the above scoring rule definition, not all of them encourage honesty and accuracy when assigning a prediction to an event. Those that do enable a fair probabilistic assignment are named \emph{proper scoring rules} \cite{lad}, which we describe in the sequence.

	Say a model's probabilistic prediction for $X$ is $P^*=(P_1^*,P_2^*,P_3^*).$ A proper scoring rule $S$ is a scoring rule  such that the mean score value 
	$$E_{P^*}[S(X,P)]=\sum_{x=1}^3 S(x,P)P^*_x$$
	 is minimized when $P=P^*$.
	In other words, if one announces
	$P$ as his probabilistic prediction
	and $S$ as a scoring rule, the lowest
	expected penalty is obtained by reporting the correct model's  uncertainty about $X$, $P^*$.
	Thus, the use of a proper scoring rule encourages the forecaster to announce $P^*$ (the correct one)
	as his probabilistic prediction  rather than some other quantity. 
	%Moreover, this method automatically leads to an overall comparison between the outcomes of different personal evaluations concerned with the same totality of events and, therefore, the accumulated score may be considered a concrete measure of success of each model.

@@

	Let $P=(P_1,P_2,P_3)$ be a probabilistic prediction  for $X$.
	The Brier score for a given outcome $x\in\{1,2,3\}$ is given by
	$$S(x,P)= \sum_{i=1}^3\mathbb{I}(x=i)(1- P_i)^2+\sum_{i=1}^3\mathbb{I}(x\neq i)P^2_i,$$
	where $\mathbb{I}$ is the indicator function.

@@

			\end{tikzpicture}
			\caption{Bi-dimensional simplex (green surface): area of possible forecasts}\label{fig:simplex}
		\end{center}
	\end{figure}

	We interpret the Brier score in the case where one of three mutually exclusive outcomes happen, the case of a football match.
	The green surface in Figure \ref{fig:simplex} represents the 2-simplex, i.e. the set of points such that $p_1+p_2+p_3=1$ for non-negative values of $p_1$, $p_2$ and $p_3$.
	The triangle representing the simplex has sides of length $\sqrt{2}$ and its height is $\sqrt{6}/2$.
	Drawing a similar equilateral triangle with height 1 and sides $2\sqrt{3}/3$, it is possible to represent all points of the simplex.
	This new triangle is used to graphically display the forecast as a internal point because the sum of the distances of every point inside it to each side, representing the probability of each event, is always one.

@@

	The Brier score for the probabilistic prediction $P=(0.25,0.35,0.40)$
	assuming the home team loses, is therefore given by $d^2=(0-0.25)^2+(0-0.35)^2+(1-0.40)^2=0.545$. 
	On the other hand, the prediction $P=(0,0,1)$ achieves score  zero, the minimum for this rule.

	It is useful to consider the score of what we will call {\it trivial prediction}:
	$P=(1/3,1/3,1/3)$.
	This assessment will produce a Brier score of $2/3$, no matter what is the final result of the match, providing, thus, a threshold that a good model should consistently beat, meaning, for the Brier score, that the scores of its predictions should be smaller than $0.667$.

@@

	The logarithmic  score is given by
	$$S(x,P)=- \sum_{i=1}^3\mathbb{I}(x=i)\ln(P_i),$$

@@

	The logarithmic score for the prediction 
	$P=(0.25,0.35,0.40)$
		 when the home team loses is therefore 
	$-\ln(0.4)\approx 0.91$. 
	On the other hand, the prediction $P=(0,0,1)$ achieves score zero, once again the minimum of this rule.
	Moreover, for the logarithmic score, the trivial prediction gives a score of approximately $1.098$.

@@

	The spherical score is given by
	$$S(x,P)=- \frac{1}{\sqrt{\sum_{i=1}^3 P^2_i}}\sum_{i=1}^3\mathbb{I}(x=x_i)P_i,$$

@@

	The spherical score for the prediction 
		$P=(0.25,0.35,0.40)$ assuming the home team loses, is given by
	$-0.4/\sqrt{0.25^2+0.35^2+0.40^2} \approx -0.68$.
	On the other hand, the prediction $P=(0,0,1)$ achieves score  $-1$ instead and, for this rule, the trivial prediction results in a score of approximately $-0.577$.

@@

	The first one is the proportion of errors made by the model or assessor. This is simply the proportion of mistakes made when considering the highest probability assessment.
	More precisely, the proportion of errors of a a sequence of probabilistic predictions for $n$ games, $P^{(1)},\ldots,P^{(n)}$
	with
		$P^{(j)}=(P^{(j)}_1,P^{(j)}_2,P^{(j)}_3)$, is defined by
	$$\frac{1}{n}\sum_{j=1}^n \mathbb{I}\left(X_j \neq \arg \max_{x \in \{1,2,3\}} P^{(j)}_x\right),$$
	where $X_j$ is the outcome of the $j$-th match.

@@

	%De Finetti used what is called the Brier score, which is just the (squared) Euclidean distance ? i.e., the sum of the squared differences between credences and (the 0/1 numerical correlates of) truth values.

@@

		\bibitem{Rue2000}
		Rue, H.; Salvesen, O. Prediction and Retrospective Analysis of
		Soccer Matches in a League. {\em Journal of the Royal Statistical
			Society - Series D} {\bf 2000}, {\em 49}, 399-418.

@@

=======

>>>>

%  LaTeX support: latex@mdpi.com
%  In case you need support, please attach any log files that you could have, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).

@@

\usepackage{changes}
\definechangesauthor[name={Rafael}, color=blue]{raf}
\setremarkmarkup{(#2)}
\definechangesauthor[name={Luis}, color= magenta]{luis}
\setremarkmarkup{(#2)}

@@

This paper is organized as follows.
Section \ref{sec::experimental} describes the proposed models, Section \ref{sec::scoring} presents the scoring rules and criteria used in this work to classify the models.
Section \ref{sec::results} reports the performance of the models predicting 190 matches comprising the second round of the Brazilian championship of first division.
In Section \ref{sec::remarks} we discuss the results and close with proposals for future research.

@@

In this section we present the probabilistic justification for the
three proposed models and the predictive distribution of an upcoming
match of team $A$ (home team) versus $B$ (away team). The main idea
is to combine the past performance of both teams in different ways
in order to predict the outcome of the match. The predictions of the
two benchmark models were published before each matchday at
\url{http://chancedegol.uol.com.br} and
\url{http://www.previsaoesportiva.com.br}.

@@

We adopt a Bayesian approach to calculate the prediction probabilities of an upcoming match of team $T$ ($A$ or $B$) based on its past performance, i.e, the number of matches it has won, tied and lost.

Let us consider the outcome of each match of team $T$ as a categorical random quantity $X$ that may assume only the values $1$, $2$, $3$, where $1$ means win ($W$), $2$ draw ($D$) and $3$ loss ($L$).
Denoting by $\theta_1, \theta_2$ and $\theta_3$ ($\theta_3 = 1-\theta_1 - \theta_2$), the probabilities of win, draw and loss, respectively, the probability mass function of $X$ is
\[
P(X=x | \boldsymbol{\theta}) = \theta_1^{\mathbb{I}_{\{1\}}(x)}
\theta_2^{\mathbb{I}_{\{2\}}(x)}(1 - \theta_1 -
\theta_2)^{{\mathbb{I}_{\{3\}}}(x)}, \qquad x \in \mathcal{X},
\]

\noindent
where $\mathcal{X}=\{1,2,3\}$ is the support of $X$,
$\mathbb{I}_{\{i\}}(x)$ is the indicator function that assumes the
value 1 if $x$ equals $i$ and 0 otherwise, and $\boldsymbol{\theta}
= (\theta_1, \theta_2)$ belogns to $\boldsymbol{\Theta} =
\{(\theta_1,\theta_2)\in [0,1]^2: \theta_1+\theta_2 \leq 1 \}$, the 2-simplex.

Assuming that $n$ matches of team $T$, given $\boldsymbol{\theta}$, are i.i.d. quantities with the above categorical distribution, and denoting by $M_1$, $M_2$ and $M_3$ the number of matches won, tied or lost by team $T$, the random vector $(M_1, M_2, M_2)$ has multinomial distribution with parameters $n$ and $\boldsymbol{\theta}$ given by
\[
P(M_1=n_1,M_2=n_2,M_3=n_3| n, \boldsymbol{\theta})=
\frac{n!}{n_1!n_2!(n-n_1-n_2)!}\theta_1^{n_1}\theta_2^{n_2}(1-\theta_1-\theta_2)^{n-n_1-n_2},
\]
\noindent 
where $n_1 + n_2 + n_3 = n$.

Our goal is to compute the predictive posterior distribution of the
upcoming match, $X_{n+1}$, that is,
$P(X_{n+1}=x|M_1=n_1,M_2=n_2,M_3=n_3)$, $x\in\mathcal{X}$. Using
Bayesian inference this probability can be easily computed as
follows. Suppose that $\boldsymbol{\theta}$ has Dirichlet prior distribution
with parameter $(\alpha_1,\alpha_2,\alpha_3)$, denoted
$\mathcal{D}(\alpha_1,\alpha_2,\alpha_3)$, with density function
\[
\pi(\boldsymbol{\theta}|\alpha)=\frac{\Gamma(\alpha_1+\alpha_2+\alpha_3)}{\Gamma(\alpha_1)\Gamma(\alpha_2)\Gamma(\alpha_3)}\theta_1^{\alpha_1-1}\theta_2^{\alpha_2-1}(1-\theta_1-\theta_2)^{\alpha_3-1}
\]
\noindent for $\alpha_1$, $\alpha_2$, $\alpha_3 > 0$, then the
posterior distribution of $\boldsymbol{\theta}$ is
$\mathcal{D}(n_1+\alpha_1,n_2+\alpha_2,n_3+\alpha_3)$. Thus, the
predictive distribution of $X_{n + 1}$ is given by the
integral
$$
P(X_{n + 1} = x | M_1 = n_1, M_2 = n_2, M_3 = n_3) = \int_{\boldsymbol{\Theta}} P(X_{n
+ 1} = x | \boldsymbol{\theta}) \pi(\boldsymbol{\theta} | M_1 = n_1, M_2 = n_2, M_3
= n_3) d\boldsymbol{\theta},
$$
which leads to the following probabilities of win, tie and loss:
\begin{align*}
P(X_{n+1} = 1 | M_1=n_1,M_2=n_2,M_3=n_3) &=
\frac{n_1+\alpha_1}{n+\alpha}\\
& \\
P(X_{n+1} = 2 | M_1=n_1,M_2=n_2,M_3=n_3) &=
\frac{n_2+\alpha_1}{n+\alpha} \\
& \\
P(X_{n+1} = 3 | M_1=n_1, M_2=n_2, M_3=n_3) &=
\frac{n_3+\alpha_1}{n+\alpha}
\end{align*}
\noindent where $\alpha=\alpha_1+\alpha_2+\alpha_3$.

In this work we used the uniform prior on the 2-simplex, i.e. $\mathcal{D}(1,1,1)$. 
This is not a restrictive prior because, since we predicted only the second round of the championship, the counts of the first round were used to update this prior---thanks to the conjugacy property of Dirichlet priors---and we may say that the actual prior we used is therefore the number of matches won, tied or lost by each team added by one.
If we had started our predictions in the first round, this prior would not have given 
a good predictive performance, but in this case we could have used team rankings such the ones provided by federations and leagues to assess a more realistic prior for each team instead.

Before we explain and illustrate the proposed models with an example, we make two further remarks.
The first one is that we will separately consider home and away games for each team. This will allow us to take into account the different performances under those conditions.
The second remark is that the prediction of an upcoming match between $A$ and $B$ can be done using
the past performance of either team $A$ or $B$.
To solve this problem, a compromise between the two predictive distributions was adopted. 
The next subsections \ref{sec::Mn_Dir1}, \ref{sec::Mn_Dir2} and \ref{sec::Mn_Dir3} present three different possible solutions.

@@

Thus, considering $X_{n + 1}$ the random outcome of this match with
respect to the home team (Gr\^{e}mio), the predictive probabilities
of $X_{n + 1}$ is obtained by equally weighting the two predictive
distributions, resulting
\begin{align*}
P(X_{n+1}=1|{\bf h}, {\bf a}) &=
\frac{1}{2}\left(\frac{h_1+\alpha_1}{h+\alpha}\right)+\frac{1}{2}\left(\frac{a_3+\alpha_3}{a+\alpha}\right)=0.5
\\
P(X_{n+1}=2|{\bf h}, {\bf a}) &=
\frac{1}{2}\left(\frac{h_2+\alpha_2}{h+\alpha}\right)+\frac{1}{2}\left(\frac{a_2+\alpha_2}{a+\alpha}\right)\simeq0.2917, \\
P(X_{n+1}=3|{\bf h}, {\bf a}) &= \frac{1}{2}
\left(\frac{h_3+\alpha_3}{h+\alpha}\right)+\frac{1}{2}\left(\frac{a_1+\alpha_1}{a+\alpha}\right)\simeq0.2083.
\end{align*}
\noindent where $h=h_1+h_2+h_3$ and $a=a_1+a_2+a_3$.

@@

The model $Mn-Dir_2$ is very similar to model $Mn-Dir_1$, but the
weights used to mix the posterior distributions are not the same.
Model $Mn-Dir_2$ weights the predictive distributions according to
the team ranks just before the match is played. In the illustrative
example, we know that after 19 matchdays, Gr\^emio was the $6^{th}$
place and Atl\'etico-PR the $11^{th}$. Therefore, we shall consider
the weights
\[w_1 = \frac{1/6}{1/6+1/11}=0.647 ~ ~ ~ ~ ~\text{and} ~ ~ ~ w_2 = \frac{1/11}{1/6+1/11}=0.353\]
\noindent for the Gr\^emio and Atl\'etico predicitive distributions,
respectively. Thus, the winning probability for Gr\^emio is
calculated as
\[P(X_{n+1}= 1|{\bf h}, {\bf a})=0.647\left(\frac{h_1+\alpha_1}{h+\alpha}\right)+0.353\left(\frac{a_3+\alpha_3}{a+\alpha}\right)=0.525,\]
\noindent and the other probabilities are computed similarly.

\subsection{Model 3: Multinomial-Dirichlet 3}
\label{sec::Mn_Dir3}

@@

Considering the example above, the combined vector ${\bf c}$ is given by ${\bf c}=(c_1,c_2,c_3)=(10,5,3)$, which leads to a posterior $\mathcal{D}(11,6,4)$ when considering the prior $\mathcal{D}(1,1,1)$.
The probability that Gr\^emio wins is therefore	
\[P(X_{n+1}=1|{\bf c})=\frac{c_1+\alpha_1}{c+\alpha}\simeq0.524
\]

\noindent
where $c=c_1+c_2+c_3$. The probabilities of drawing and losing are computed similarly.

@@

How can we fairly distinguish better-informed states of knowledge from more poorly informed states?
A way of doing this is to use proper scoring rules for evaluating states of uncertain knowledge, where
the score may be interpreted as a numerical measure of how inaccurate was the prevision.

%More precisely, a scoring rule is a function of two variables (or vectors, matrices etc) that assigns a real number (score) to each possible outcome $X$ and its respective prevision value, $P(X)$.
%It is maximum for each $x$ when $x=P(X)$, and is nonincreasing both as $x$ departs from $P(X)$ (for each $P(X)$) and as $P(X)$ departs from $x$ (for each $x$).
%Thus, the best score is achieved when knowledge of $X$ is asserted as the value of $X$ itself, whatever it might be. Most of the scoring rules also exhibit an arbitrary normalizing feature such as $S[x,P(X)]=0$ for each $x$.

%Although many functions satisfy the criteria for a scoring rule specified above, not every scoring rule can be regarded as providing a fair assessment of the assertion $P(X)$, that is, not all scoring rules encourage honesty and accuracy in assessing numerical values for their previsions.
%Those that do encourage a fair assessment are named \emph{proper scoring rules} \cite{lad}.

%Say a forecaster's uncertainty about $X$ is $P_0$. A proper scoring rule is a scoring rule $S$ such that the prevision value $P(X)$ that minimizes $E_{P_0}[S(X,P(X))]$ is $P_0$.
%In other words, if one receives a reward $S(x,P(x))$ when $x$ happens and if $S$ is a proper scoring rule, then the highest expected reward is obtained by reporting one's  true uncertainty about $X$.
%The use of a proper scoring rule thus encourages the forecaster to be honest to maximize the expected reward.
%Moreover, this method leads automatically to an overall comparison between the outcomes of different personal evaluations concerned with the same totality of events.
%Therefore, the accumulated score may be considered a concrete measure of success of each forecaster or model.

More precisely, a scoring rule is a function of two variables (or vectors, matrices etc) that assigns a real number (score) to each possible outcome of $X$ and its respective prevision value, $P_X$, i.e. $P_X$ is the probability assigned to each possible outcome of $X$.
It is minimum for each $x$ when $P_X$ is such that $P_X(X=x)=1$, and is nondecreasing both as $x$ departs from $P_X$ (for each $P_X$) and as $P_X$ departs from $x$ (for each $x$).
Thus, the best score (the smaller value of the score) is achieved when knowledge of $X$ is asserted as the value of $X$ itself, whatever it might be\footnote{A scoring rule can also be defined to be such that a \emph{large} value of the score indicates good predictions.}.
%Most of the scoring rules also exhibit an arbitrary normalizing feature such as $S(x,P_X)=0$ for each $x$.

Although many functions satisfy the criteria for a scoring rule specified above, not every scoring rule can be regarded as providing a fair assessment of the assertion $P_X$, that is, not all scoring rules encourage honesty and accuracy in assessing numerical values for their previsions.
Those that do encourage a fair assessment are named \emph{proper scoring rules} \cite{lad}.

Say a forecaster's uncertainty about $X$ is $P_0$. A proper scoring rule is a scoring rule $S$ such that the prevision value $P_X$ that minimizes $E_{P_0}[S(X,P_X)]$ is $P_0$.
In other words, if one receives a penalty $S(x,P_X)$ when $x$ happens and if $S$ is a proper scoring rule, then the lowest expected penalty is obtained by reporting one's  true uncertainty about $X$.
The use of a proper scoring rule thus encourages the forecaster to be honest to minimize the expected penalty.
Moreover, this method leads automatically to an overall comparison between the outcomes of different personal evaluations concerned with the same totality of events and, therefore, the accumulated score may be considered a concrete measure of success of each forecaster or model.

@@

Let $P_X$ be a probability assignment for $X$.
The Brier score is given by
$$S(X;P_X)= \sum_{i}\mathbb{I}(X=x_i)(1- P_X(x_i))^2+\sum_{i}\mathbb{I}(X\neq x_i)P_X^2(x_i),$$
where $\mathbb{I}$ is the indicator function.

@@

\end{tikzpicture}
\caption{Bi-dimensional simplex (green surface): area of possible predictions}\label{fig:simplex}
\end{center}
\end{figure}

We interpret the Brier score in the case where one of three mutually exclusive outcomes happen, the case of a football match.
The green surface in Figure \ref{fig:simplex} represents the 2-simplex, i.e. the set of points such that $p_1+p_2+p_3=1$ for non-negative values of $p_1$, $p_2$ and $p_3$.
The triangle representing the simplex has sides of length $\sqrt{2}$ and its height is $\sqrt{6}/2$.
Drawing a similar equilateral triangle with height 1 and sides $2\sqrt{3}/3$, it is possible to represent all points of the simplex.
This new triangle is used to graphically display the predictions as internal points because the sum of the distances of every point inside it to each side, representing the probability of each event, is always one.

@@

The Brier score for the prediction $P_X=(0.25,0.35,0.40)$, assuming the home team loses, is therefore given by $d^2=(0-0.25)^2+(0-0.35)^2+(1-0.40)^2=0.545$. 
If the prediction was $P_X=(0,0,1)$, the score would be zero, the minimum for this rule.

It is useful to consider the score of what we will call {\it trivial prediction}: $(1/3,1/3,1/3)$.
This assessment will produce a Brier score of $2/3$, no matter what is the final result of the match, providing, thus, a threshold that a good model should consistently beat, meaning, for the Brier score, that the scores of its predictions should be smaller than $0.667$.

@@

The logarithmic  score is given by
$$S(X;P_X)=- \sum_{i}\mathbb{I}(X=x_i)\ln(P_X(x_i)),$$

@@

The logarithmic score for the prediction $P_X=(0.25,0.35,0.40)$, when the home team loses is therefore 
$-\ln(0.4)\approx 0.91$. 
If the prediction was $P_X=(0,0,1)$, the score would be zero, once again the minimum of this rule.
For the logarithmic score, the trivial prediction gives a score of approximately $1.098$.

@@

The spherical score is given by
$$S(X;P_X)=- \frac{1}{\sqrt{\sum_i P_X^2(x_i)}}\sum_{i}\mathbb{I}(X=x_i)P_X(x_i),$$

@@

The spherical score for the prediction $P_X=(0.25,0.35,0.40)$, assuming the home team loses, is given by
$-0.4/\sqrt{0.25^2+0.35^2+0.40^2} \approx -0.68$.
If the prediction was $P_X=(0,0,1)$, the score would be $-1$ instead and, for this rule, the trivial prediction results in a score of approximately $-0.577$.

@@

The first one is the proportion of errors made by the model or assessor. This is simply the proportion of mistakes made when considering the highest probability assessment.
More precisely, the proportion of errors of a given prediction $P_X$ is defined by
$$\frac{1}{n}\sum_{j=1}^n \mathbb{I}\left(X_j \neq \arg \max_{x} P_X(x_{j})\right),$$
where $X_j$ is the outcome of the $j$-th match and $P_X(x_{j})$ is the probability attributed to result $x\in\mathcal{X}=\{1,2,3\}$ for the $j$-th match.

@@

%De Finetti used what is called the Brier score, which is just the (squared) Euclidean distance — i.e., the sum of the squared differences between credences and (the 0/1 numerical correlates of) truth values.

@@

\bibitem{Rue2000}
Rue, H.; Salvesen, O. Prediction and Retrospective Analysis of
Soccer Matches in a League {\em Journal of the Royal Statistical
Society - Series D} {\bf 2000}, {\em 49}, 399-418.

@@
